\chapter{Research Plan} \label{ch:research_plan}

\section{Proposal} \label{sec:proposal}

\todo[inline]{Possible to remove the accuracy part from the proposal}

\begin{bluebox}{Research Proposal}
    \noindent To investigate the effectiveness of an artificial viscosity term in improving the stability of online, data-driven \gls{les}-\gls{vms} simulations applied to the one-dimensional Burgers equation, using reinforcement learning.
\end{bluebox}

From the proposal, the primary research question states:

\begin{enumerate}[label=\textbf{\color{titlecolor}PRQ:}]
    \item \textit{To what extent can an artificial viscosity term\todo{name it term or corrector?} improve the stability and accuracy of VMS-ANN closure models applied to the one-dimensional Burgers equation?}
\end{enumerate}

With subsequent research questions:

\begin{enumerate}[label=\textbf{\color{titlecolor}RQ\arabic*.}]

%    \item \textbf{Question 1}
%    \begin{enumerate}[label=\textbf{\color{titlecolor}SQ\arabic{enumi}.\alph*so}]
%        \item Subquestion 1.1
%        \item
%    \end{enumerate}

    \item \textit{How does the \gls{av} term affect numerical stability across a range of coarse-grid resolutions compared to a no-\gls{av} baseline model?}

    \item \textit{How does the artificial viscosity term affect the energy dissipation and resolved energy evolution of the \gls{vms}-\gls{ann} model compared to theoretical dissipation levels?}

    \item \textit{How does the introduction of the artificial viscosity term affect the accuracy of the \gls{vms}-\gls{ann} model compared to a no-\gls{av} baseline model?}
        \begin{enumerate}[label=\textbf{\color{titlecolor}SQ\arabic{enumi}.\alph*}]
        \item \textit{How does the artificial viscosity term affect the correlation of the resolved flow solution with \gls{dns} data, \textbf{a posteriori}?} \todo{a posteriori can be removed as we already make the distinction the AV term acts on online simulations}
        \item \textit{Does the artificial viscosity corrector modify the input distribution observed by the fixed closure \gls{mlp}, and does this lead to observable differences in closure predictions relative to the no-\gls{av} case?}
    \end{enumerate}

    \item \textit{Does a locally applied \gls{av} term perform better in terms of accuracy and stability compared to a global one?}
        \begin{enumerate}[label=\textbf{\color{titlecolor}SQ\arabic{enumi}.\alph*}]
            \item \textit{Does a local model achieve better-resolved flow correlations with DNS data than the global model?}
            \item \textit{Can a local model perform longer, numerically stable simulations than the global model?}
        \end{enumerate}
\end{enumerate}

\section{Project Plan} \label{sec:plan}

Task list:

\begin{enumerate}[label=\textbf{\color{titlecolor}\arabic*.}]
    \item Create 1D \gls{fem} solver - 1.5 weeks
    \begin{enumerate}[label=\textbf{\color{titlecolor}\arabic{enumi}.\alph*}]
        \item Develop solver code
        \item Verify solver
    \end{enumerate}

    \item Generate training data - 1.5 weeks
    \begin{enumerate}[label=\textbf{\color{titlecolor}\arabic{enumi}.\alph*}]
        \item Run \gls{dns}
        \item Extract subgrid closures (project \gls{dns} data)
    \end{enumerate}

    \item Train fixed \gls{mlp} model - 1 weeks

    \item Evaluate \emph{a priori} correlations between fixed model and \gls{dns} - 0.5 weeks

    \item Train \gls{gav} corrector model - 2 weeks

    \item Run \gls{les}-\gls{gav} model - 0.5 weeks

    \item Evaluate stability and accuracy - 1 week

    \item[] Prepare Midterm - $1$ week

    \item[] 9 weeks → Midterm meeting

    \item Train \gls{lav} corrector model - 2 weeks

    \item Run \gls{lav} model - 0.5 weeks

    \item Evaluate stability and accuracy - 1 week

    \item[] Answer Research Questions

    \item Run \gls{gav}, \gls{lav} across multiple grid sizes - 1 week

    \item Evaluate numerical stability across multiple grid sizes - 1 week

    \item Evaluate energy behavior - 1 week

    \item Evaluate accuracy and correlations with \gls{dns} - 1 week

    \item Evaluate differences between \gls{gav}, \gls{lav} and fixed \gls{mlp} - 2 week

    \item Check accuracy differences between \gls{gav} and \gls{lav} - 1 week

    \item Check stability duration between \gls{gav} and \gls{lav} - 1 week

    \item[] 8 weeks → If successful, attempt 3D problem.

    \item[] If applicable:

    \item Train model for 3D environment - $3$ weeks

    \item Evaluate performance of 3D problem

    \item Investigate differences between 1D and 3D problems

\end{enumerate}
